---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# FastForecast

<!-- badges: start -->

<!-- badges: end -->

The goal of FastForecast is to provide a fast and accurate forecasting method for time series data. This package takes a **dataframe** of several time series as input, and can correct for **atypical points**, **seasonality** and **stationarity**. It then generates forecasting models using **6 econometric models** and **6 Machine-Learning models**. It also allows you to display series graphically, calculate forecast quality indicators and display them in table form as well as graphically.

## Installation

You can install the development version of FastForecast from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("NoaLRX/FastForecast")
```

## Functions 

Here are the different functions that you can use in this package.

### Time-Series transformation

```{r, eval=FALSE}
library(FastForecast)
ts_transfo <- function(DATAFRAME, YEAR, MONTH, FREQUENCY)
```

This function transforms the $n$ columns of a dataframe into $n$ individuals Time Series.

Each time series will take the name `"ts_"` followed by the name of the column in the dataframe.

To use this function you need to pass a dataframe as the first argument, a year, month and frequency as the last argument.

[For example:]{.underline}

```{r, eval=FALSE}
ts_transfo(land_w, 2013, 01, 4) # Start the 1st January 2013, with quarterly data
ts_transfo(land_w, 1996, 05, 12) # Start the 5 May 1996, with monthly data
```

### Atypical Points Correction

This function takes no argument, and will **correct** your previously created time-series and correct any atypical points. This function use the `tso()` function from the package [tsoutliers](https://www.rdocumentation.org/packages/tsoutliers/versions/0.6-8/topics/tso).

```{r, eval=FALSE}
library(tsoutliers)
atypical_tso()
```

### Seasonality Detection

This function takes no argument, and will **detect** any seasonality on your previously created time-series. This function use the `combined_test()` and `seasdum()` functions from the package [seastests](https://www.rdocumentation.org/packages/seastests/versions/0.15.4).

```{r, eval=FALSE}
library(seastests)
seaso_detect()
```

### Seasonality Correction

This function takes no argument, and will **correct** any seasonality on your previously created time-series. This function use the `stl()` function from the package [stats](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/stl).

```{r, eval=FALSE}
library(stats)
seaso_correct()
```

### Seasonality Verification

This function takes no argument, and will **detect** any seasonality on your previously created and corrected time-series. This function use the `combined_test()` and `seasdum()` functions from the package [seastests](https://www.rdocumentation.org/packages/seastests/versions/0.15.4).

The main purpose of this function is to ensure that there is no seasonality in the corrected series.

```{r, eval=FALSE}
library(seastests)
seaso_verif()
```

### Stationarity Treatment

This function takes no argument, and will **detect** and **correct,** if necessary, the time series that are not stationary. The function detect stationarity with the `adf.test()` function from the package [tseries](https://www.rdocumentation.org/packages/tseries/versions/0.10-54/topics/adf.test).

```{r, eval=FALSE}
library(tseries)
statio()
```

If after a first differentiation, the time-serie is still not stationary, the function will differentiate a second time. If after that, the time serie is still not stationary, the function will print a message and stop any treatment.

### Aligning Time-Series length

This function takes no argument, and will **align** your corrected time series that may have different length, because of differentiations.

```{r, eval=FALSE}
allign_ts()
```

### Creating a dataframe with each time-series

This function takes one argument, the $Y_t$ variable, the one that you're trying to predict.

This function will create a single data frame `results_df` with each time series as columns, with the $Y_t$ as the first variable.

```{r, eval=FALSE}
create_df("my_Y_variable")
```

### Variables Selection

This function uses two methods of variables selections. The first is the ***"BestSubSet"*** method from the [leaps](https://cran.r-project.org/web/packages/leaps/leaps.pdf) package. The second is the ***"Gets"*** method from the [gets](https://cran.r-project.org/web/packages/gets/gets.pdf) package.

This function takes one argument, the $Y_t$ variable, the one that you're trying to predict.

```{r, eval=FALSE}
vselec("my_Y_variable")
```

After using the function, you should modify your `results_df` dataframe and remove the variables not retained by the two methods.

### Econometrics Forecasting Models

[This function takes three arguments:]{.underline}

-   `DATAFRAME`: Your dataframes with $n$ rows and $p$ columns (with $n > p$)

-   `Y_VARIABLE`: Your $Y_t$ variable, meaning the variable you're trying to predict.

-   `PERIOD`: The frequency of your time-series *(12 for monthly, 4 for quarterly...)*

```{r, eval=FALSE}
eco_models(df, "My_Y_Variable", 12)
eco_models(dataframe, "Gas_Price", 4)
```

[This function uses the following models and packages:]{.underline}

-   ARX model with gets ([gets](https://cran.r-project.org/web/packages/gets/gets.pdf))

-   ARX moddel with auto.arima ([forecast](https://www.rdocumentation.org/packages/forecast/versions/8.21.1))

-   ARMAX model with auto.arima ([forecast](https://www.rdocumentation.org/packages/forecast/versions/8.21.1))

-   Naive model ([forecast](https://www.rdocumentation.org/packages/forecast/versions/8.21.1))

-   LM model ([stats](https://www.rdocumentation.org/packages/stats/versions/3.6.2))

-   AR1 mode ([forecast](https://www.rdocumentation.org/packages/forecast/versions/8.21.1))

-   GAM model ([mgcv](https://www.rdocumentation.org/packages/mgcv/versions/1.9-0))

### Machine-Learning Forecasting Models

[This function takes three arguments:]{.underline}

-   `DATAFRAME`: Your dataframes with $n$ rows and $p$ columns (with $n > p$)

-   `Y_VARIABLE`: Your $Y_t$ variable, meaning the variable you're trying to predict.

```{r, eval=FALSE}
ml_models(df, "My_Y_Variable")
```

[This function uses the following models and packages:]{.underline}

-   MLP models ([neuralnet](https://www.rdocumentation.org/packages/neuralnet/versions/1.44.2))

-   MARS model ([earth](https://www.rdocumentation.org/packages/earth/versions/5.3.2))

-   SVM model ([e1071](https://www.rdocumentation.org/packages/e1071/versions/1.1-3))

-   Random Forest ([randomForest](https://www.rdocumentation.org/packages/randomForest/versions/4.7-1.1))

-   XGB Boost ([xgboost](https://www.rdocumentation.org/packages/xgboost/versions/1.7.5.1))

-   kNN ([caret](https://www.rdocumentation.org/packages/caret/versions/6.0-94))
